DEBS Grand Challenge: A High-Throughput, Scalable Solution for Calculating Frequent Routes and Profitability of New York Taxis

### 2015-05-19 presentation/grand_challenge.tex
updating the presentation

\documentclass[]{beamer}

@@

\author[Amila Suriarachchi]{Amila Suriarachchi \\ Shrideep Pallickara } % Your name
\institute[CSU] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Colorado State University \\ % Your institution for the title page
\medskip
\textit{amilas@cs.colostate.edu} % Your email address
\medskip
}
\date{\today} % Date, can be changed to a custom date

@@

\begin{frame}
\frametitle{Contributions}
\begin{itemize}
        \item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity
        \item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity 
	\item We have developed a scalable parallel evaluation technique to process queries efficiently
\end{itemize}

@@

\begin{frame}
\frametitle{NodeList Data Structure}

@@

\begin{figure}
        \includegraphics[width=0.8\linewidth]{orderedlist.png}
\end{figure}

@@

\begin{figure}
        \includegraphics[width=0.7\linewidth]{DinamicHeap.png}
\end{figure}

@@

\begin{frame}
\frametitle{Runing Median (1/2)}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{runningmediang.png}
\end{figure}

@@

\begin{frame}
\frametitle{Runing Median (2/2)}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{runningmediand.png}
\end{figure}

@@

\begin{algorithm}[H]
\footnotesize
before $\gets$ nodeList.getTopValues()\;
window.add(event)\;
\eIf{ nodeList.containsKey(event.route) }{
        routCount $\gets$ nodeList.get(event.route)\;
        routCount.count++\;
        nodeList.decrementPosition(event.route)\;
}{
        nodeList.add(event.route, new RouteCount(1))\;
}

\While{ there exists expired events }{
        expEvent $\gets$ window.poll()\;
        routCount $\gets$ nodeList.get(expEvent.route)\;
        routCount.count--\;
        \eIf{ routCount.count is 0 }{
                nodelist.remove(expEvent.route)\;
        }{
                nodeList.incrementPosition(expEvent.route)\;
        }
}

@@

\begin{frame}
\frametitle{Profitable Cells Query}
\begin{itemize}
        \item Median fare calculation
		\begin{itemize}
			\item Use the PriorityQueue based running median algorithm
		\end{itemize}
        \item Handling empty taxis
		\begin{itemize}
                        \item Generate two events (Pickup event and Drop event) for each dataset event
                \end{itemize}
        \item Handling drop event expiration
		\begin{itemize}
                        \item Keep taxi ids within the profitable cell object
                \end{itemize}
\end{itemize}

@@

\begin{figure}
        \includegraphics[width=0.8\linewidth]{sequential.png}
\end{figure}

@@

>>>>

\documentclass[handout]{beamer}

@@

\author[Amila Suriarachchi]{Amila Suriarachchi \\ Shrideep Pallickara } % Your name
\institute[CSU] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Colorado State University \\ % Your institution for the title page
\medskip
\textit{\{amilas,shrideep\}@cs.colostate.edu} % Your email address
\medskip
}
\date{\today} % Date, can be changed to a custom date

@@

%\begin{frame}
%\frametitle{Contributions}
%\begin{itemize}
%        \item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity
%        \item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity 
%	\item We have developed a scalable parallel evaluation technique to process queries efficiently
%\end{itemize}
%\end{frame}

\begin{frame}
\frametitle{NodeList Data Structure (1/2)}

\begin{itemize}
        \item Provides notion of a sorted list of values starting from position 0
        \item Values can be accessed using the key (eg. Route, Cell)
\end{itemize}

@@

\begin{frame}
\frametitle{NodeList Data Structure (2/2)}

\begin{itemize}
        \item \textit{OrderedList} data structure is used to keep first ten values
        \item \textit{DynamicHeap} data structure is used to keep other values
\end{itemize}

@@

\begin{itemize}
        \item A Doubly linked List is used to store values
        \item A Map is used to retrieve \textit{ListNode} objects with O(1) time
\end{itemize}

\begin{figure}
        \includegraphics[width=0.7\linewidth]{orderedlist.png}
\end{figure}

@@

\begin{itemize}
        \item A Heap is used to store values
        \item \textit{HeapNode} keeps the position of the array to remove, moveUp and moveDown objects with O(log n) time
	\item A Map is used to retrieve \textit{HeapNode} objects with O(1) time
\end{itemize}

\begin{figure}
        \includegraphics[width=0.6\linewidth]{DinamicHeap.png}
\end{figure}

@@

\begin{frame}
\frametitle{Running Median (1/2)}

\begin{itemize}
        \item Two heaps are used to keep values greater than the current median and less than the current median
        \item A \textit{PriorityQueue} based implementation
		\begin{itemize}
        		\item Takes O(log n) time to add a value 
        		\item Takes O(n) time to remove a value
		\end{itemize}
\end{itemize}

\begin{figure}
        \includegraphics[width=0.6\linewidth]{runningmediang.png}
\end{figure}

@@

\begin{frame}
\frametitle{Running Median (2/2)}

\begin{itemize}
        \item Two heaps are used to keep values greater than the current median and less than the current median
        \item A \textit{DynamicHeap} based implementation with value as the heap key and count as the heap value 
                \begin{itemize}
                        \item Takes O(log n) time to add and remove a value
                \end{itemize}
\end{itemize}

\begin{figure}
        \includegraphics[width=0.6\linewidth]{runningmediand.png}
\end{figure}

@@

\begin{algorithm}[H]
\footnotesize
before $\gets$ nodeList.getTopValues()\;
window.add(event)\;
\eIf{ nodeList.containsKey(event.route) }{
        routeCount $\gets$ nodeList.get(event.route)\;
        routeCount.count++\;
        nodeList.decrementPosition(event.route)\;
}{
        nodeList.add(event.route, new RouteCount(1))\;
}

\While{ there exists expired events }{
        expEvent $\gets$ window.poll()\;
        routeCount $\gets$ nodeList.get(expEvent.route)\;
        routeCount.count--\;
        \eIf{ routeCount.count is 0 }{
                nodelist.remove(expEvent.route)\;
        }{
                nodeList.incrementPosition(expEvent.route)\;
        }
}

@@

\begin{frame}
\frametitle{Profitable Cells Query}
\begin{itemize}
        \item Median fare calculation
		\begin{itemize}
			\item Use the \textit{PriorityQueue} based running median algorithm
		\end{itemize}
        \item Handling empty taxis
		\begin{itemize}
                        \item Generate two events (Pickup event and Drop event) for each dataset event
                \end{itemize}
        \item Handling drop event expiration
		\begin{itemize}
                        \item Keep taxi ids within the profitable cell object
                \end{itemize}
\end{itemize}

@@

\begin{itemize}
        \item One \textit{QueryEvaluator} instance is used
        \item \textit{EventEmitter} and \textit{QueryEvaluator} can be executed in,
                \begin{itemize}
                        \item A single JVM using one thread
                        \item A single JVM using different threads with queues to communicate
                        \item A distributed setup with TCP to communicate 
                \end{itemize}
\end{itemize}

\begin{figure}
        \includegraphics[width=0.7\linewidth]{sequential.png}
\end{figure}

@@

\begin{itemize}
        \item Multiple \textit{QueryEvaluator} instances are used
        \item \textit{EventEmitter} and \textit{QueryEvaluators} can be executed in,
                \begin{itemize}
                        \item A Single JVM using different threads with queues to communicate
                        \item A distributed setup with TCP to communicate
                \end{itemize}
\end{itemize}
